Microlensing as a possible probe of the internal structure of quasars

### 2015-07-17 paper1/fitting.tex
fitting changes

In this section we simulate the fitting of source models to
lightcurves in the presence of random and systematic errors.  We
generate mock data for three source models --- a crescent, a uniform
disc and a Gaussian disc --- and then fit each mock light curve to all
three source models. Parameter fitting and marginalisation was done by
Markov-chain Monte Carlo (MCMC).

\begin{figure}
\centering
\includegraphics[width=0.9\hsize,bb=0 0 576 432
                ]{plots/data_cc.eps}
\includegraphics[width=0.9\hsize,bb=0 0 576 432
                ]{plots/data_dd.eps}
\includegraphics[width=0.9\hsize,bb=0 0 576 432
                ]{plots/data_gg.eps}
\caption{\label{fig:mockdata} Three different mock datasets with
  three different parametrisations fitted to each --- crescent (red),
  uniform-disk (blue) and Gaussian disk (green).}
\end{figure}

@@

The crescent source had parameters (as explained in
\S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
and inner circles and $(a,b)$ being the coordinates of the inner
circle with respect to the centre of the outer circle.  The parameter
values were
\begin{equation}
   (R_p, R_n, a, b) = (50.0, 30.0, 15.0, 10.0).
\label{eqn:cp}
\end{equation}
The uniform disc had the same (outer) radius as the crescent.  In the
Gaussian source, we set $3\sigma=50$.  For convenience, below we will
refer to $R_p$ of the Gaussian source, by which we actually mean
$3\sigma$.

Each mock light curve also has three nuisance parameters, namely the
beginning and end of the event and the brightness normalization.
These are to be marginalised out by the MCMC.

Figure~\ref{fig:mockdata} shows the three light curves.  Each has
250 points regularly spaced in time, with Gaussian noise at the level
of 10\% of the current brightness.  The effective signal to noise can
be considered to be $0.1\times\sqrt{250}\simeq150$.

@@

Each of the three mock light curves was fitted to all three models.
We denote the nine possible cases with letter pairs, with the first
letter denoting the model fitted and the second letter representing
the source: thus CG means a {\em C\/}rescent model was fitted to mock
data from a {\em G\/}aussian source, DC stands for a uniform {\em
  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
and so on.

@@

Let us first consider the three cases where a crescent model was
fitted.  These are the solid curves in Figure~\ref{fig:mcmc}, with the
colours of the curves indicating the source.  Meanwhile,
Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
values.

@@

\item[2) {\bf CD}:] solid blue curve and blue ellipses.  Here a
  crescent model was fitted to a light curve from a uniform disc.  The
  situation is formally a crescent with $R_n=0$ and $a,b$ arbitrary,
  and this shows in the blue ellipses in the recovered parameter
  values.  The redundant parameters $a$ and $b$, in effect, allow the
  model to partly fit the noise, and hence the $\chi^2$ is somewhat
  lower than in the previous case.

@@

Next we consider the three cases where a disc model is fitted.  These
correspond to he dashed curves in Figure~\ref{fig:mcmc}.  There is
only one interesting parameter to fit, the disc radius $R_p$, and no
equivalent of Figure~\ref{fig:crescentfit} is needed.

@@

Finally, we consider the three cases where a Gaussian model is fitted.
These correspond to he dotted curves in Figure~\ref{fig:mcmc}.  Again,
there is only one interesting parameter to fit, the Gaussian
$3\sigma$-radius, which we have called $R_p$.

@@

The above results suggest the following strategy for fitting a
lightcurve from an unknown source profile: first try a Gaussian-source
model; if the data reject that model, try a uniform disc; if the
uniform disc is also rejected by the data, try a crescent model.  Our
numerical experiments indicate --- assuming one of the three source
models is correct --- that the reduced $\chi^2$ would unmask the
correct model, and its parameters would be correctly recovered.

>>>>

In this section we simulate the fitting of source models to
lightcurves in the presence of random and systematic errors.  We
generate mock data for three source models --- a crescent, a uniform
disc and a Gaussian disc --- and then fit each mock light curve to all
three source models. Parameter fitting and marginalising was done by
Markov-chain Monte Carlo (MCMC).

\begin{figure}
\centering
\includegraphics[width=0.9\hsize,bb=0 0 576 432
                ]{plots/data_cc.eps}
\includegraphics[width=0.9\hsize,bb=0 0 576 432
                ]{plots/data_dd.eps}
\includegraphics[width=0.9\hsize,bb=0 0 576 432
                ]{plots/data_gg.eps}
\caption{\label{fig:mockdata} Three different mock datasets with
  three different parametrizations fitted to each --- crescent (red),
  uniform-disk (blue) and Gaussian disk (green).}
\end{figure}

@@

The crescent source had parameters (as explained in
\S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
and inner circles and $(\alpha,\beta)$ being the coordinates of the inner
circle with respect to the centre of the outer circle in the coordinate system associated with
the image in figure 1 and not associated with the caustic surface as defined in the previous sections.  The parameter values were
\begin{equation}
   (R_p, R_n, $\alpha$, $\beta$) = (50.0, 30.0, 15.0, 10.0).
\label{eqn:cp}
\end{equation}
The uniform disc had the same (outer) radius as the crescent.  In the
Gaussian source, we set $3\sigma=50$.  For convenience, below we will
refer to $R_p$ of the Gaussian source, by which we actually mean
$3\sigma$.

Each mock light curve also has three nuisance parameters, namely the
beginning and end of the event and the brightness normalisation.
These are to be marginalised out by the MCMC.

Figure~\ref{fig:mockdata} shows the three light curves.  Each has
250 points regularly spaced in time, with Gaussian noise at the level
of 10\% of the current brightness.  The effective signal to noise can
be considered to be \textbf{ bad math here -- $0.1\times\sqrt{250}\simeq150$}.

@@

Each of the three mock light curves was fitted to all three models.
We denote the nine possible cases with letter pairs, with the first
letter denoting the assumed model for the fitting procedure and 
the second letter representing
the source: thus CG means a {\em C\/}rescent model was fitted to mock
data from a {\em G\/}aussian source, DC stands for a uniform {\em
  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
and so on.

@@

Let us first consider the three cases where a crescent model was
fitted.  These are the solid curves in Figure~\ref{fig:mcmc}, with the
colours of the curves indicating the source.  Meanwhile,
Figure~\ref{fig:crescentfit} shows $2\sigma$ of the inferred parameter
values. \textbf{In the following enumerated conclusions we associate the word "wrong"
with parameters for which the real value corresponds to a PDF of the fitted model smaller than 
0.01. }

@@

\item[2) {\bf CD}:] solid blue curve and blue ellipses.  Here a
  crescent model was fitted to a light curve from a uniform disc.  The
  situation is formally a crescent with $R_n=0$ and $a,b$ arbitrary,
  and this shows in the blue ellipses in the recovered parameter
  values.  The redundant parameters $\alpha$ and $\beta$, in effect, allow the
  model to partly fit the noise, and hence the $\chi^2$ is somewhat
  lower than in the previous case.

@@

Next we consider the three cases where a disc model is fitted.  These
correspond to the dashed curves in Figure~\ref{fig:mcmc}.  There is
only one interesting parameter to fit, the disc radius $R_p$, and no
equivalent of Figure~\ref{fig:crescentfit} is needed.

@@

Finally, we consider the three cases where a Gaussian model is fitted.
These correspond to the dotted curves in Figure~\ref{fig:mcmc}.  Again,
there is only one interesting parameter to fit, the Gaussian
$3\sigma$-radius, which we have called $R_p$.

@@

The above results suggest the following strategy for fitting a
lightcurve from an unknown source profile: first try a Gaussian-source
model; if the data reject that model, try a uniform disc; if the
uniform disc is also rejected by the data, try a crescent model.  Our
numerical experiments indicate --- assuming one of the three source
models is correct --- that the reduced $\chi^2$ would unmask the
correct model, \textbf{and its parameters would be correctly recovered}.
